Changing Dynamics of Urban Bike Thefts: Understanding Trends to Inform Future Prevention Efforts

### 2024-12-03 paper/paper.qmd
Modified some details in the final version

The measurement process for obtaining bicycle theft data involves transforming real-world events into a structured dataset, such as the Toronto Police Service's bicycle theft records. This dataset is a product of the process by which bicycle thefts are reported to the police and is primarily obtained through online reporting tools or direct communication with law enforcement.

The reporting process for bicycle thefts under 5,000 dollars requires the victim to provide as much detailed information as possible, including the serial number, customizations (e.g., unique colors or features), and the estimated value of the bicycle. This information is important to ensure the accuracy and traceability of individual cases in the dataset. Bicycles reported must have a value of less than 5,000 dollars or they will be reported through a different channel that differentiates between minor and major thefts.

Once a report is submitted, it is reviewed by the appropriate authorities and, if verified, the incident becomes an entry in the dataset used for analysis. Researchers can then analyze this data to identify patterns and trends, examining variables such as time of theft and location types. This methodical approach effectively links real-world theft events with their analytical representations, enabling a thorough examination of the dynamics and risk factors tied to different locations and premises within the dataset. This structured data transformation process is essential for generating ideas about bicycle theft, providing a foundation for strategies to mitigate these incidents effectively. These attributes help to understand the context of each burglary and allow for further analysis to identify burglary patterns or high-risk areas.

@@

```{r}
#| label: fig-hour-bike-thefts
#| fig-cap: "Time-Specific Patterns in Bike Thefts: Hourly Distribution Analysis"
#| echo: false
#| warning: false
#| message: false

@@

The Bayesian logistic regression model predicts the likelihood of a bike theft ($y_i$), where ($y_i = 1$) indicates a theft and ($y_i = 0$) otherwise. The model includes several predictors: $X_{\text{Occurrence Hour},i}$ captures the hour of the day (0–23) to identify temporal patterns in theft likelihood, and $X_{\text{Occurrence Month},i}$ represents the month (1–12) to account for seasonal variations. $X_{\text{Premises Type},i}$ categorizes the theft locations into "Apartment" (the baseline category), "House", "Outdoors", and "Other", examining how location influences theft probability. Additionally, $X_{\text{Bike Cost},i}$ is a continuous variable representing the bike's reported value, hypothesizing that higher values may increase theft likelihood due to perceived value.

Specifically, $\beta_1$ relates to $X_{\text{Occurrence Hour},i}$, exploring how different hours affect the log-odds of theft, while $\beta_2$ for $X_{\text{Occurrence Month},i}$ examines how theft likelihood fluctuates monthly. The coefficients $\beta_3$ for $X_{\text{Premises Type},i}$ show the variation in theft log-odds for "House", "Outdoors", and "Other" compared to the baseline "Apartment". For example, a positive $\beta_3$ for "House" would imply higher log-odds of theft in house settings compared to apartments. Lastly, $\beta_4$ for $X_{\text{Bike Cost},i}$ quantifies how the bike's cost impacts theft risk, suggesting that more expensive bikes are more likely targets.

These $\beta$ coefficients are essential for understanding how variables such as $X_{\text{Occurrence Hour},i}$, $X_{\text{Occurrence Month},i}$, and $X_{\text{Premises Type},i}$ influence the likelihood of a reported bike theft. Analyzing these relationships helps pinpoint specific patterns in theft occurrences, showing whether certain times of day, types of premises, or bike cost ranges significantly alter theft risk.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

The Bayesian logistic regression model was chosen based on prior knowledge about factors influencing bike theft. For instance, premises type and residential status are expected to have different theft risks based on environmental characteristics, while bike cost reflects the value of the stolen item, which might influence the likelihood of theft. Bayesian regression allows us to incorporate this prior understanding through weakly informative priors, enhancing the model's ability to capture these relationships. Additionally, variables like occurrence hour and premises type may exhibit complex patterns that the Bayesian approach can better account for by combining prior knowledge with observed data. The model provides not only point estimates but also posterior distributions, enabling a clearer understanding of uncertainty in each predictor's impact. However, the model assumes a linear relationship between predictors and the log-odds of theft, as well as independence between theft events. However, these assumptions may be limiting. For example, if theft events are highly correlated spatially or temporally—such as clusters of thefts occurring in the same area or during specific times—the model may not accurately capture these patterns.

The model comparison indicates that both the Bayesian logistic regression model and the frequentist logistic regression model achieved identical accuracy scores of 98.57% on the test dataset. This high accuracy suggests that both models effectively captured the relationship between the predictor variables (e.g., Occurrence_Hour, Occurrence_Month, Premises_Type, Bike_Cost) and the binary outcome (Theft_Status). The process of comparing these models involved splitting the dataset into training (80%) and testing (20%) subsets to evaluate their predictive performance on unseen data. Both models were trained on the training set, with their predictions compared against the actual values in the test set. Classifications were made using a threshold of 0.5 for predicted probabilities, and accuracy was computed as the proportion of correctly classified cases. These results demonstrate that both modeling approaches are highly effective for this particular dataset, yielding near-identical predictive performance.

@@

## Methodology Framework
### Introduction to the Bike Theft Analysis Methodology

@@

### Survey and Sampling Methodology

@@

### Recruitment Process

@@

## Data Collection and Analysis
### Data Collection Protocol

@@

### Statistical Analysis of Survey and Police Data

@@

## Implementation and Community Engagement
### Budget Allocation and Resource Management

@@

### Community Engagement and Feedback Integration

@@

## Reporting and Conclusion
### Transparency and Reporting

@@

### Conclusion

@@

# Data details {#sec-data-details}

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: "Preview of Cleaned Bike Theft Data "
#| echo: false
#| warning: false
#| message: false

@@

## Posterior prediction check
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior prediction check"
Bike_Theft_model <-
  readRDS(file = here::here("models/Bike_Theft_model.rds"))

>>>>

The measurement process for obtaining bicycle theft data involves transforming real-world events into a structured dataset, such as the Toronto Police Service's bicycle theft records. This dataset is a product of the process by which bicycle thefts are reported to the police and is primarily obtained through online reporting tools or direct communication with law enforcement. The reporting process for bicycle thefts under 5,000 dollars requires the victim to provide as much detailed information as possible, including the serial number, customizations (e.g., unique colors or features), and the estimated value of the bicycle. This information is important to ensure the accuracy and traceability of individual cases in the dataset. Bicycles reported must have a value of less than 5,000 dollars or they will be reported through a different channel that differentiates between minor and major thefts. Once a report is submitted, it is reviewed by the appropriate authorities and, if verified, the incident becomes an entry in the dataset used for analysis. Researchers can then analyze this data to identify patterns and trends, examining variables such as time of theft and location types. This methodical approach effectively links real-world theft events with their analytical representations, enabling a thorough examination of the dynamics and risk factors tied to different locations and premises within the dataset,  providing a foundation for strategies to mitigate these incidents effectively.

@@

```{r}
#| label: fig-hour-bike-thefts
#| fig-cap: "Time-Specific Patterns in Bike Thefts: Hourly Distribution Plot"
#| echo: false
#| warning: false
#| message: false

@@

The Bayesian logistic regression model predicts the likelihood of a bike theft **($y_i$)**, where **($y_i = 1$)** indicates a theft and **($y_i = 0$)** otherwise. The model includes several predictors:

- **$X_{\text{Occurrence Hour},i}$** :Represents the hour of the day, ranging from 0 to 23. This variable is used to identify temporal patterns that may affect the likelihood of bike theft at different times of the day.
- **$X_{\text{Occurrence Month},i}$** :Indicates the month, numbered from 1 to 12.

- **$X_{\text{Premises Type},i}$** : Categorizes the location of the theft into one of several types: "Apartment" (used as the baseline category), "House", "Outdoors", and "Other".

- **$X_{\text{Bike Cost},i}$** :A continuous variable that represents the reported monetary value of the bicycle.

Specifically, **$\beta_1$** relates to **$X_{\text{Occurrence Hour},i}$**, exploring how different hours affect the log-odds of theft, while **$\beta_2$** for **$X_{\text{Occurrence Month},i}$** examines how theft likelihood fluctuates monthly. The coefficients $\beta_3$ for **$X_{\text{Premises Type},i}$** show the variation in theft log-odds for "House", "Outdoors", and "Other" compared to the baseline "Apartment". For example, a positive **$\beta_3$** for "House" would imply higher log-odds of theft in house settings compared to apartments. Lastly, **$\beta_4$** for **$X_{\text{Bike Cost},i}$** quantifies how the bike's cost impacts theft risk, suggesting that more expensive bikes are more likely targets. Analyzing these relationships helps pinpoint specific patterns in theft occurrences, showing whether certain times of day, types of premises, or bike cost ranges significantly alter theft risk.

We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm. We use the default priors from `rstanarm`.

@@

The Bayesian logistic regression model was chosen based on prior knowledge about factors influencing bike theft. This model helps us understand how different factors like where the bike is and how much it costs might make theft more likely. We use priors in the model, which means we start with an initial guess based on previous knowledge and then update this as we get more data. This method is good for figuring out complex behaviors from the data, like when thefts happen more often or in certain places. The model doesn’t just give us a single number; it shows a range of possible outcomes, which helps us understand the uncertainty or reliability of our predictions. However, the model assumes that each theft is independent and that the relationship between the factors we study and theft is straightforward. This could be a problem if, for example, thefts happen in clusters or at specific times and places, which our model might not pick up accurately.

The model comparison shows that both the Bayesian logistic regression model and the frequentist logistic regression model achieved identical accuracy scores of 98.57% on the test dataset. This high accuracy suggests that both models effectively captured the relationship between the predictor variables (e.g., Occurrence_Hour, Occurrence_Month, Premises_Type, Bike_Cost) and the binary outcome (Theft_Status). The process of comparing these models involved splitting the dataset into training (80%) and testing (20%) subsets to evaluate their predictive performance on unseen data. Both models were trained on the training set, with their predictions compared against the actual values in the test set. Classifications were made using a threshold of 0.5 for predicted probabilities, and accuracy was computed as the proportion of correctly classified cases. These results demonstrate that both modeling approaches are highly effective for this particular dataset, yielding near-identical predictive performance.

@@

## Introduction to the Bike Theft Analysis Methodology

@@

## Survey and Sampling Methodology

@@

## Recruitment Process

@@

## Data Collection Protocol

@@

## Statistical Analysis of Survey and Police Data

@@

## Budget Allocation and Resource Management

@@

## Community Engagement and Feedback Integration

@@

## Transparency and Reporting

@@

## Conclusion

@@

# Data details {#sec-data-details}
@tbl-cleaned-data shows the preview of the Cleaned Bike Theft Dataset.
```{r}
#| label: tbl-cleaned-data
#| tbl-cap: "Preview of Cleaned Bike Theft Data "
#| echo: false
#| warning: false
#| message: false

@@

## Posterior prediction check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior prediction check"
Bike_Theft_model <-
  readRDS(file = here::here("models/Bike_Theft_model.rds"))
